---
title: "Central data monitoring report"
subtitle: "Coordinating centre view — Synthetic Adaptive Platform Trial in Critical Illness"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    embed-resources: true
    fig-width: 9
    fig-height: 4.5
    code-fold: true
    code-summary: "Show the code behind this section"
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
suppressMessages({
  library(dplyr)
  library(ggplot2)
  library(targets)
})

ROOT <- normalizePath("..", winslash = "/")
for (file in list.files(file.path(ROOT, "R"), pattern = "[.]R$",
                        recursive = TRUE, full.names = TRUE)) {
  source(file)
}
store <- file.path(ROOT, "_targets")

cfg <- load_trial_config()
sites <- resolve_sites(cfg)

forms        <- tar_read(conformed_forms, store = store)
findings     <- tar_read(findings, store = store)
endpoint     <- tar_read(dawols, store = store)
recall       <- tar_read(recall_table, store = store)
recall_total <- tar_read(recall_summary, store = store)
conformance  <- tar_read(conformance_log, store = store)
unmatched    <- tar_read(unmatched_findings, store = store)

completeness <- tar_read(monthly_completeness, store = store)
timeliness   <- tar_read(monthly_timeliness, store = store)
drift        <- tar_read(delay_drift, store = store)
attention    <- tar_read(attention_list, store = store)

house_theme <- theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold", size = 12),
        plot.subtitle = element_text(colour = "grey35", size = 10),
        legend.position = "bottom")

participants <- n_distinct(forms$randomisation$participant_id)
```

::: {.callout-note appearance="simple"}
`r participants` participants · `r nrow(forms$randomisation)` randomisations ·
`r nrow(sites)` sites in `r length(unique(sites$country))` countries ·
rule set `r rule_set_version()` ·
generated `r format(Sys.Date(), "%d %B %Y")`
:::

# Sites needing attention

Ordered by entry-delay drift, then by critical findings. Deliberately **not** a
single composite score: a score hides which of the three problems a site has,
and the coordinating centre needs to know which call to make.

```{r attention}
top <- attention |>
  head(10) |>
  transmute(Site = site_id, Name = site_name, Country = country,
            `Critical findings` = critical_findings,
            `Endpoint records` = records,
            `Not evaluable` = not_evaluable,
            `% incomplete` = pct_incomplete,
            `Drift (days/month)` = slope_days_per_month)
knitr::kable(top)
```

# Entry-delay drift

No validation rule catches drift. Every record at a drifting site is perfectly
valid — it was simply entered late, which is not a rule violation. Only the
trend across months shows a site falling further behind, and only measuring it
against each site's **own** initiation shows that it is not merely a young site
still settling in.

```{r drift-plot}
drift_plot <- drift |>
  mutate(site_id = factor(site_id, levels = rev(site_id)),
         flagged = slope_days_per_month > 0.5)

ggplot(drift_plot, aes(site_id, slope_days_per_month, fill = flagged)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values = c("FALSE" = "#4C72B0", "TRUE" = "#C44E52"),
                    labels = c("Within expectation", "Investigate"), name = NULL) +
  labs(title = "Change in median data entry delay, per month since site initiation",
       subtitle = "Positive means the site is getting slower over time",
       x = NULL, y = "Days added per month") +
  house_theme
```

```{r drift-detail}
worst <- drift$site_id[1]
worst_monthly <- timeliness |>
  filter(site_id %in% c(worst)) |>
  group_by(month) |>
  summarise(median_delay = median(median_delay_days), .groups = "drop")

ggplot(worst_monthly, aes(month, median_delay)) +
  geom_line(colour = "#C44E52", linewidth = 0.9) +
  geom_point(colour = "#C44E52", size = 2) +
  scale_x_date(date_labels = "%b %Y") +
  labs(title = paste("Median entry delay at", worst),
       subtitle = paste0("The site with the steepest drift: ",
                         drift$first_median_days[1], " days rising to ",
                         drift$last_median_days[1], " days"),
       x = NULL, y = "Median days from event to entry") +
  house_theme
```

# Findings across the trial

```{r findings-overall}
by_rule <- findings_by_rule(findings)
knitr::kable(
  by_rule |>
    transmute(Rule = rule_id, Name = rule_name, Severity = severity,
              Form = form, Findings = findings, Participants = participants,
              Sites = sites)
)
```

```{r findings-by-site-plot}
by_site <- findings |>
  filter(severity == "critical") |>
  count(site_id, name = "critical") |>
  right_join(sites[, c("site_id", "country")], by = "site_id") |>
  mutate(critical = coalesce(critical, 0L)) |>
  arrange(desc(critical))

ggplot(by_site, aes(reorder(site_id, critical), critical, fill = country)) +
  geom_col() +
  coord_flip() +
  labs(title = "Critical findings by site",
       subtitle = "Raw counts. Larger sites will have more; read alongside the per-site reports.",
       x = NULL, y = "Critical findings", fill = NULL) +
  house_theme
```

# How well do the rules actually work?

The validation engine is scored against a catalogue of defects deliberately
injected into the data, so its performance is measured rather than asserted.

```{r recall}
recall_display <- recall |>
  transmute(Defect = defect_id, Name = defect_name,
            `Expected rule` = expected_rule_id,
            Injected = injected, Detected = detected,
            Recall = ifelse(is.na(recall), "n/a", sprintf("%.1f%%", 100 * recall)))
knitr::kable(recall_display)
```

**Overall: `r sprintf("%.1f%%", 100 * recall_total$recall)`** —
`r recall_total$detected` of `r recall_total$scoreable_by_a_rule` defect records
that a rule targets.

A further **`r recall_total$not_scoreable_monitoring_signals` records** belong to
defect types that **no rule in this milestone targets**. They are reported as
`n/a` rather than as 0%, because scoring a rule set against defects it was never
written to catch would misrepresent it, and omitting them would misrepresent the
defect catalogue. These are the statistical signals — entry drift, terminal-digit
preference, adverse-event under-reporting — and the drift section above is the
first of them to be addressed.

## Findings not traceable to an injected defect

Not the same as false positives. A finding with no matching injected defect is
usually a genuine problem the simulation created incidentally, or a real
ambiguity in the data model.

```{r unmatched}
knitr::kable(
  unmatched |>
    arrange(desc(not_traceable)) |>
    transmute(Rule = rule_id, Name = rule_name, Severity = severity,
              Findings = findings,
              `Traceable to injection` = traceable_to_injected_defect,
              `Not traceable` = not_traceable)
)
```

`LOG-001` is the clearest example. No form in this trial carries an ICU
discharge date, so a gap in daily records caused by discharge and readmission is
indistinguishable from a gap caused by missing data entry. Querying both is what
a data manager would actually do, and the fix is a schema change rather than a
rule change.

# What the ingest layer had to undo

Every transformation applied between the site's export and the analysis dataset
is logged. This table is the evidence of what was done to the data.

```{r conformance}
knitr::kable(
  conformance_summary(conformance) |>
    transmute(Transformation = transformation, Sites = sites,
              `Log entries` = rows, `Values affected` = values)
)
```

# Primary endpoint completeness by site

A day with no daily record counts as **unknown**, never as a day free of life
support. Missing records therefore understate the endpoint. The percentage
incomplete is a direct measure of how much each site's data is degrading the
outcome the trial exists to measure.

```{r endpoint-by-site}
endpoint_site <- endpoint |>
  group_by(site_id) |>
  summarise(records = n(),
            not_evaluable = sum(is.na(days_alive_without_life_support)),
            pct_incomplete = 100 * mean(!complete),
            .groups = "drop") |>
  arrange(desc(pct_incomplete))

ggplot(endpoint_site, aes(reorder(site_id, pct_incomplete), pct_incomplete)) +
  geom_col(fill = "#4C72B0") +
  coord_flip() +
  labs(title = "Proportion of endpoint records with at least one unknown day",
       subtitle = "Every unknown day understates days alive without life support",
       x = NULL, y = "% incomplete") +
  house_theme
```

```{r endpoint-overall}
knitr::kable(summarise_dawols(endpoint))
```

The mean over all records and the mean over complete records differ. That
difference is not a rounding detail: it is larger than the one-day margin the
analysis plan uses to declare two arms practically equivalent, which is why the
endpoint reports its own completeness rather than returning a single number.

---

::: {.callout-warning appearance="simple"}
All data in this report is **synthetic**, generated by the code in this
repository from a fixed seed. Site names are invented. This is an independent
educational project with no affiliation to any real trial, hospital or research
group.
:::
