---
title: "Site data monitoring report"
subtitle: "Synthetic Adaptive Platform Trial in Critical Illness"
params:
  site_id: "DK-01"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    embed-resources: true
    fig-width: 9
    fig-height: 4
    code-fold: true
    code-summary: "Show the code behind this section"
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
suppressMessages({
  library(dplyr)
  library(ggplot2)
  library(targets)
})

# The report renders from reports/, so everything is resolved against the
# project root rather than the working directory.
ROOT <- normalizePath("..", winslash = "/")
for (file in list.files(file.path(ROOT, "R"), pattern = "[.]R$",
                        recursive = TRUE, full.names = TRUE)) {
  source(file)
}
store <- file.path(ROOT, "_targets")

site_id <- params$site_id
cfg <- load_trial_config()
sites <- resolve_sites(cfg)
site <- sites[sites$site_id == site_id, ]
stopifnot(nrow(site) == 1)

forms      <- tar_read(conformed_forms, store = store)
findings   <- tar_read(findings, store = store)
endpoint   <- tar_read(dawols, store = store)

# Precomputed by the pipeline: every site report needs the all-site picture to
# compare against, and recomputing it per report is the same work 25 times.
completeness <- tar_read(monthly_completeness, store = store)
timeliness   <- tar_read(monthly_timeliness, store = store)

site_completeness <- completeness |> filter(site_id == !!site_id)
site_timeliness   <- timeliness   |> filter(site_id == !!site_id)
site_finding_rows <- site_findings_detail(findings, site_id)
site_endpoint     <- endpoint |> filter(site_id == !!site_id)

# "Recent" means the last six months of data this site actually has, not the
# last six calendar months, so a site that has paused enrolment still gets a
# readable trend rather than an empty chart.
recent_months <- sort(unique(site_completeness$month))
recent_months <- tail(recent_months, 6)
recent_completeness <- site_completeness |> filter(month %in% recent_months)

months_open <- floor(as.numeric(Sys.Date() - site$initiation_date) / 30.44)

plural <- function(n, singular, plural_form = paste0(singular, "s")) {
  if (is.na(n)) return("")
  paste(n, if (n == 1) singular else plural_form)
}

house_theme <- theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold", size = 12),
        plot.subtitle = element_text(colour = "grey35", size = 10),
        legend.position = "bottom")
```

::: {.callout-note appearance="simple"}
**`r site$site_name`** (`r site_id`), `r site$country_name`.
Initiated `r format(site$initiation_date, "%d %B %Y")`, `r months_open` months ago.
Report generated `r format(Sys.Date(), "%d %B %Y")` from data version
`r if (nrow(findings)) findings$data_version[1] else "working"`,
rule set `r rule_set_version()`.
:::

# What to do this week

```{r actions}
actions <- site_actions(site_finding_rows, site_endpoint, site_completeness)
```

`r if (!nrow(actions)) "**Nothing outstanding.** No critical findings, no unevaluable endpoints, and no falling completeness trend at this site."`

```{r actions-list, results="asis"}
if (nrow(actions)) {
  for (i in seq_len(nrow(actions))) {
    count <- if (is.na(actions$records[i])) "" else
      paste0(" - **", plural(actions$records[i], "record"), "**")
    cat(sprintf("%d. **%s**%s  \n    %s\n\n",
                actions$priority[i], actions$action[i], count, actions$why[i]))
  }
}
```

> This list is the report. Everything below it is the evidence behind these
> items, and exists so that a query can be answered with "here is why we are
> asking" rather than "the system flagged it".

# Open findings

```{r findings-summary}
if (nrow(site_finding_rows)) {
  summary_table <- site_finding_rows |>
    group_by(severity, rule_id, rule_name) |>
    summarise(findings = n(), participants = n_distinct(participant_id),
              .groups = "drop") |>
    arrange(severity, desc(findings))
  knitr::kable(summary_table, col.names = c("Severity", "Rule", "Name",
                                           "Findings", "Participants"))
} else {
  cat("No open findings at this site.")
}
```

## Participants to follow up

Critical findings only, with the identifiers needed to act on them.

```{r findings-detail}
critical <- site_finding_rows |> filter(severity == "critical")
if (nrow(critical)) {
  detail <- critical |>
    select(rule_id, participant_id, domain, form, field, observed_value) |>
    arrange(rule_id, participant_id) |>
    head(50)
  knitr::kable(detail, col.names = c("Rule", "Participant", "Domain", "Form",
                                     "Field", "Observed value"))
  if (nrow(critical) > 50) {
    cat("\n\n*Showing the first 50 of ", nrow(critical), " critical findings.*", sep = "")
  }
} else {
  cat("No critical findings at this site.")
}
```

# Completeness

Proportion of required, human-entered fields carrying a value, by the month of
the clinical event. Identifiers and system-stamped dates are excluded: they are
always present and would flatter the measure.

```{r completeness-trend}
if (nrow(recent_completeness)) {
  ggplot(recent_completeness, aes(month, completeness, colour = form)) +
    geom_line(linewidth = 0.7) +
    geom_point(size = 1.8) +
    scale_y_continuous(labels = scales::percent, limits = c(NA, 1)) +
    scale_x_date(date_labels = "%b %Y") +
    labs(title = "Completeness over the last six months of activity",
         subtitle = "A falling line is a process problem; a low flat line is a backlog",
         x = NULL, y = "Fields complete", colour = NULL) +
    house_theme
} else {
  cat("No completeness data for this site.")
}
```

```{r completeness-table}
if (nrow(recent_completeness)) {
  table_data <- recent_completeness |>
    mutate(month = format(month, "%b %Y"),
           completeness = sprintf("%.1f%%", 100 * completeness)) |>
    select(form, month, records, completeness) |>
    tidyr::pivot_wider(names_from = month, values_from = completeness,
                       id_cols = form)
  knitr::kable(table_data)
}
```

# Timeliness

Days from the clinical event to entry into the EDC. Read this against the
holiday column: entry slows around local public holidays in every country, and
a report that ignores that generates queries nobody can act on.

```{r timeliness-trend}
site_monthly_delay <- site_timeliness |>
  group_by(month) |>
  summarise(median_delay = median(median_delay_days),
            p90 = max(p90_delay_days),
            holiday_affected = max(holiday_affected), .groups = "drop")

if (nrow(site_monthly_delay)) {
  ggplot(site_monthly_delay, aes(month, median_delay)) +
    geom_col(aes(fill = holiday_affected > 0.15), width = 22) +
    geom_line(aes(y = p90), colour = "grey30", linetype = "dashed") +
    scale_fill_manual(values = c("FALSE" = "#4C72B0", "TRUE" = "#DD8452"),
                      labels = c("Ordinary month", "Public holidays in month"),
                      name = NULL) +
    scale_x_date(date_labels = "%b %Y") +
    labs(title = "Median days from event to data entry",
         subtitle = "Dashed line is the 90th percentile. A rising trend is the signal, not the level.",
         x = NULL, y = "Days") +
    house_theme
} else {
  cat("No timeliness data for this site.")
}
```

# This site against all sites

Compared at the same **stage of participation**, not on a calendar axis. A site
in its third month is compared against every other site's third month. Comparing
a young site against an established one on calendar time penalises it for being
young, which helps nobody.

```{r vs-all-completeness}
comparison <- compare_to_all_sites(completeness, sites, site_id, "completeness")
comparison <- comparison |> filter(months_since_initiation <= 24)

if (any(!is.na(comparison$this_site))) {
  ggplot(comparison, aes(months_since_initiation)) +
    geom_ribbon(aes(ymin = all_sites_lower, ymax = all_sites_upper),
                fill = "grey85") +
    geom_line(aes(y = all_sites_median), colour = "grey45", linewidth = 0.6) +
    geom_line(aes(y = this_site), colour = "#C44E52", linewidth = 0.9) +
    geom_point(aes(y = this_site), colour = "#C44E52", size = 1.8) +
    scale_y_continuous(labels = scales::percent) +
    labs(title = paste("Completeness at", site_id, "against all sites"),
         subtitle = "Red is this site. Grey band is the interquartile range across all sites.",
         x = "Months since site initiation", y = "Fields complete") +
    house_theme
} else {
  cat("Not enough data to compare this site.")
}
```

```{r vs-all-timeliness}
delay_comparison <- compare_to_all_sites(
  timeliness |> group_by(site_id, month) |>
    summarise(median_delay_days = median(median_delay_days), .groups = "drop"),
  sites, site_id, "median_delay_days")
delay_comparison <- delay_comparison |> filter(months_since_initiation <= 24)

if (any(!is.na(delay_comparison$this_site))) {
  ggplot(delay_comparison, aes(months_since_initiation)) +
    geom_ribbon(aes(ymin = all_sites_lower, ymax = all_sites_upper),
                fill = "grey85") +
    geom_line(aes(y = all_sites_median), colour = "grey45", linewidth = 0.6) +
    geom_line(aes(y = this_site), colour = "#C44E52", linewidth = 0.9) +
    geom_point(aes(y = this_site), colour = "#C44E52", size = 1.8) +
    labs(title = paste("Entry delay at", site_id, "against all sites"),
         subtitle = "Red is this site. A line that climbs while the band stays flat is drift.",
         x = "Months since site initiation", y = "Median days to entry") +
    house_theme
} else {
  cat("Not enough data to compare this site.")
}
```

# Primary endpoint completeness

Days alive without life support at 30 days is the outcome every adaptive
analysis reads. A day with no daily record counts as **unknown**, not as a day
free of life support, so missing records understate the endpoint rather than
inflating it.

```{r endpoint}
if (nrow(site_endpoint)) {
  endpoint_summary <- summarise_dawols(site_endpoint)
  knitr::kable(
    data.frame(
      Measure = c("Participant-domain records", "Not evaluable (no randomisation date)",
                  "Complete", "Incomplete (some days unknown)",
                  "Died within 30 days", "Total unknown days",
                  "Mean, all records", "Mean, complete records only"),
      Value = c(endpoint_summary$n, endpoint_summary$not_evaluable,
                endpoint_summary$n_complete, endpoint_summary$n_incomplete,
                endpoint_summary$died_within_window,
                endpoint_summary$total_unknown_days,
                endpoint_summary$mean_all, endpoint_summary$mean_complete_only)
    )
  )
} else {
  cat("No endpoint records for this site yet.")
}
```

---

::: {.callout-warning appearance="simple"}
All data in this report is **synthetic**, generated by the code in this
repository from a fixed seed. The site names are invented. This is an
independent educational project with no affiliation to any real trial,
hospital or research group.
:::
